---
title: "Breach Data Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This analysis explores data breaches, identifying trends, common exposed data types, and the impact of breaches on different services.

## Load Libraries and Dataset
```{r load_libraries, message=FALSE, warning=FALSE}
# Install missing packages
packages <- c("tidyverse", "lubridate", "knitr")
new_packages <- packages[!(packages %in% installed.packages()[,"Package"])]

if(length(new_packages)) install.packages(new_packages)

# Load libraries
library(tidyverse)
library(lubridate)
library(knitr)
```

```{r load_data}
data <- read.csv("breached_services_info.csv", stringsAsFactors = FALSE)
```

## Data Cleaning and Preparation
```{r data_cleaning}
data$BreachDate <- as.Date(data$BreachDate, format="%Y-%m-%d")
data$AddedDate <- as.Date(data$AddedDate, format="%Y-%m-%d")
data$ModifiedDate <- as.Date(data$ModifiedDate, format="%Y-%m-%d")
```

## Number of Breaches Per Year
```{r breaches_per_year}
data %>%
  mutate(Year = year(BreachDate)) %>%
  group_by(Year) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = Year, y = n)) +
  geom_col(fill = "blue") +
  theme_minimal() +
  labs(title = "Number of Breaches per Year", x = "Year", y = "Count")
```

## Top 10 Services with Highest Breaches
```{r top_services}
data %>%
  arrange(desc(PwnCount)) %>%
  select(Name, PwnCount) %>%
  head(10) %>%
  ggplot(aes(x = reorder(Name, PwnCount), y = PwnCount)) +
  geom_col(fill = "red") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Top 10 Services with Highest Breaches", x = "Service", y = "Pwned Accounts")
```

## Verified vs Unverified Breaches
```{r verified_vs_unverified}
data %>%
  group_by(IsVerified) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = IsVerified, y = n, fill = IsVerified)) +
  geom_col() +
  theme_minimal() +
  labs(title = "Verified vs Unverified Breaches", x = "Verified", y = "Count")
```

## Breach Severity Analysis
```{r breach_severity}
data <- data %>%
  mutate(BreachSeverity = case_when(
    PwnCount < 1000000 ~ "Small",
    PwnCount >= 1000000 & PwnCount < 10000000 ~ "Medium",
    PwnCount >= 10000000 ~ "Large"
  ))

data %>%
  group_by(BreachSeverity) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = BreachSeverity, y = n, fill = BreachSeverity)) +
  geom_col() +
  theme_minimal() +
  labs(title = "Breach Severity Distribution", x = "Severity", y = "Count")
```

## Trends in Sensitive Data Breaches
```{r sensitive_breaches, warning=FALSE, message=FALSE}
if ("IsSensitive" %in% colnames(data)) {
  data$IsSensitive <- as.logical(data$IsSensitive)  # Ensure correct format
  
  sensitive_breaches <- data %>%
    filter(!is.na(BreachDate) & IsSensitive == TRUE) %>%
    mutate(Year = year(BreachDate)) %>%
    group_by(Year) %>%
    summarise(n = n())
  
  if(nrow(sensitive_breaches) > 0) {
    ggplot(sensitive_breaches, aes(x = Year, y = n)) +
      geom_line(color = "darkred", size = 1.2) +
      theme_minimal() +
      labs(title = "Trends in Sensitive Data Breaches", x = "Year", y = "Count")
  } else {
    print("No data available for sensitive breaches.")
  }
} else {
  print("Column 'IsSensitive' not found in dataset.")
}
```

## Distribution of Breaches by Domain
```{r breach_by_domain}
data %>%
  group_by(Domain) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  head(10) %>%
  ggplot(aes(x = reorder(Domain, -n), y = n, fill = Domain)) +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  labs(title = "Top Domains Affected by Breaches", x = "Domain", y = "Number of Breaches")
```

## Time Between Breach and Reporting
```{r breach_reporting_time}
data %>%
  mutate(ReportDelay = as.numeric(AddedDate - BreachDate)) %>%
  ggplot(aes(x = ReportDelay)) +
  geom_histogram(binwidth = 30, fill = "orange", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Reporting Delay", x = "Days Between Breach and Report", y = "Frequency")
```

## Conclusion
This analysis provides key insights into breach trends, affected services, and exposed data. Addressing these vulnerabilities is crucial for enhancing digital security.
